Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-12-05 dsn-2018/bare_conf.tex

A \textit{cryptographic hash function} is a function:
\begin{align*}
  h: S_{i} \times B \rightarrow S_{o}
\end{align*}
which satisfies the usual properties of cryptographic hash functions
(preimage, second preimage, collision, deterministic, etc.). We denote
the input state space as $S_{i}$, the input block space as $B$, and the
output state space as $S_{o}$. Usually the input state space and the
output state space are the same, so we omit the subscripts, $S$. We
assume that these are sets of binary strings of a fixed length. In the
case of MD4, $S$ is all binary strings of length 128, and $B$ is all
binary strings of length 512.

@@

We use the term \textit{intermediate state variables} to refer to
intermediate steps in the computation of a hash function under
a specific input. Typically these are the outputs of
one-way compression functions as part of the Merkel-Damg{\aa}rd
construct. If $i$ is an input to a hash function, we use $I(i)$ to
denote the intermediate state variables. Formally, this is represented
as an ordered tuple of binary strings of the size of the updated state.
In the case of MD4, this is an ordered tuple of cardinality 48, where
each index is a binary string of length 32, for a total of 1536
intermediate state variables. We denote the space that $I(i)$ maps
into as $\mathcal{V}$. We define a set of functions $R(x)$ where $x$
is an element derived from the sets $\mathcal{C}$ or $\mathcal{F}$
which returns an element.

    \item A \textit{collision}, $L \subseteq \mathcal{I}$ is a subset of the
        input space of a hash function with $\big|L\big| \geq 2$, such that
        there exists an output $v \in S$ such that for all inputs $i \in L$,
        $h(i) = c$.

        \begin{itemize}
            \item We define a \textit{simple collision} to be any collision
                with cardinality exactly two.

            \item We define a \textit{strong collision} to be a collision which
                has multiple blocks which hash to the same output under the same
                input  state. That is, if $L$ is a collision, then
                $\forall i \in L$, $i_s = c$ for some $c \in S$ for $L$ to be
                a strong collision.

            \item We define a \textit{multicollision} to be a collision which
                has multiple input states which hash to the same output under
                a single input block. That is, if $L$ is a collision, then
                $\forall i \in L$, $i_b = c$ for some $c \ in B$ for $L$ to be
                a multicollision.
        \end{itemize}

    \item For a simple collision, we say that the \textit{differential path}
        is the difference between the two sets of intermediate state
        variables. \textbf{TODO -- HELP define formal notation?}
            \begin{itemize}
                \item Signed - Wang's, we don't use.
                \item Unsigned - more common, actually used by us.
            \end{itemize}

    \item A \textit{collision class}, $C$, is an arbitrary differential path.
        If there exists at least one collision pair which has that differential
        path, we call that collision class non-empty. We denote the set of all
        non-empty collision classes as $\mathcal{C}$. When we wish to convert
        between a collision, $c \in L$ and a collision class,
        $C \in \mathcal{C}$, we use the notation $C = \mathcal{C}(c)$.

    \item A \textit{family of collision classes}, $F$, is the indices of a
        collision class, $C$, which have any non-zero difference. That is,
        $F = \{ 0 \leq i \leq R : C[i] \neq 0 \}$.
\end{itemize}

>>>>

We use the following terminology when discussing the notion of collisions in
a hash function.

A \textit{cryptographic hash function} is a function:
\begin{align*}
  h: S_{i} \times B \rightarrow S_{o}
\end{align*}
which satisfies the usual properties of cryptographic hash functions
(preimage, second preimage, collision, deterministic, etc.). We denote
the input state space as $S_{i}$, the input block space as $B$, and the
output state space as $S_{o}$. Usually the input state space and the
output state space are the same, so we omit the subscripts, $S$. We
assume that these are sets of binary strings of a fixed length. In the
case of MD4, $S$ is all binary strings of length 128, and $B$ is all
binary strings of length 512. We assume that the hash function can be
reduced to an arbitrary number of rounds, $r$. We note the hash
function reduced to $r$ rounds as $h_{r}$.

@@

We use the term \textit{intermediate state variables} to refer to
intermediate steps in the computation of a hash function under
a specific input. Typically these are the outputs of
one-way compression functions as part of the Merkel-Damg{\aa}rd
construct. If $i$ is an input to a hash function, we use $I(i)$ to
denote the intermediate state variables. Formally, this is represented
as an ordered tuple of binary strings of the size of the updated state.
In the case of MD4, this is an ordered tuple of cardinality $r$, where
each index is a binary string of length 32, for a maximum of $1536$
intermediate state variables. We denote the space that $I(i)$ maps
into as $\mathcal{V}$.

A \textit{collision}, $L \subseteq \mathcal{I}$ is a subset of the
input space of a hash function with $\big|L\big| \geq 2$, such that
there exists an output $v \in S$ such that for all inputs $i \in L$,
$h(i) = c$. We extend this to arbitrary round hash functions using
the notation $(r, L)$, to explicitly denote that $L$ is a collision
under $h_{r}$.

\begin{itemize}
\item We define a \textit{multicollision} to be a collision which
  has multiple input states which hash to the same output under
  a single input block. That is, if $L$ is a collision, then
  $\forall i \in L$, $i_b = c$ for some $c \ in B$ for $L$ to be
  a multicollision.
\end{itemize}

For a simple collision, we say that the \textit{differential path}
is the difference between the two sets of intermediate state
variables. This is formally represented by an ordered tuple of
$r$ elements. We use the simple xor difference throughout this paper.

A \textit{collision class}, $C$, is an arbitrary differential path.
If there exists at least one collision pair which has that differential
path, we call that collision class non-empty. We denote the set of all
non-empty collision classes as $\mathcal{C}$. When we wish to convert
between a collision, $c \in L$ and a collision class,
$C \in \mathcal{C}$, we use the notation $C = \mathcal{C}(c)$. When
$c$ is a tuple, $(r, L)$, we say that the collision class, $C$
must exist over same round collisions, and thus $C$ becomes a
two-tuple, $(r, d)$, where $d$ is the ordered tuple denoting
differential path.

A \textit{family of collision classes}, $F$, is the indices of a
collision class, $C$, which have any non-zero difference. That is,
$F = \{ 0 \leq i \leq R : C[i] \neq 0 \}$. When $C$ again is over
a single round, $r$, we extend $F$ likewise, to become
$(r, F)$. We denote the set of all families of collision classes as
$\mathcal{F}$.
